---
title: "Lecture 13"
---


### Lecture handout: 
chp9-handout.pdf

### Textbook: 
Chapter 9: Multiple and Logistic Regression


### R Topics

#### generating a regression problem
```{r}
# first generate x, explanatory variable
x <- rnorm(100, mean=50, sd=25)

# set population slope and intercept
B0 <- 100
B1 <- -10

# generate error/residuals
err <- rnorm(100, mean=0, sd=100)

# finally generate y, response variable
y <- B0 + B1*x + err

# now find the sample estimates
R <- cor(x,y)
b1 <- cor(x,y)*sd(y)/sd(x)
# or
b1 <- lm(y~x)$coefficients['x']
# solve for intercept given slope and mean(x), mean(y)
b0 <- -b1*mean(x)+mean(y)
# or 
b0 <- lm(y~x)$coefficients['(Intercept)']

# predicted y values
yhat <- b0 + b1*x
# base R plot
plot(x,y)
lines(x,yhat)
# ggplot2 plot
library(ggplot2)
ggplot(data.frame(x=x,y=y,yhat=yhat)) + geom_point(aes(x=x,y=y), alpha = .2) + geom_line(aes(x=x,y=yhat))
ggplot(data.frame(x=x,y=y,yhat=yhat,err=err)) + geom_point(aes(x=x,y=y,alpha=abs(err))) + geom_line(aes(x=x,y=yhat))
ggplot(data.frame(x=x,y=y,yhat=yhat,err=err)) + geom_point(aes(x=x,y=y,alpha=abs(err))) + geom_line(aes(x=x,y=yhat)) + geom_segment(aes(x=x, y=y, xend=x, yend=yhat, alpha=abs(err) )) 

# examine SStot, SSreg, SSerr
SStot <- var(y)
SSreg <- var(yhat)
SSerr <- var(y-yhat)
SStot == SSreg + SSerr # not exact
all.equal(SStot, SSreg + SSerr) # TRUE
# R^2 equalities
all.equal(R^2, 1 - SSerr/SStot)
all.equal(R^2, SSreg/SStot)
```

#### Extra discussion on correlation and covariance
```{r}
all.equal(cor(x,y), cov(x,y)/sd(x)/sd(y))
```

recall variance 

LaTeX: $var\left(x\right) =\frac{1}{n-1}\sum_{i=1}^{n}(x_i-\bar{x})^2 $
v a r ( x ) = 1 n − 1 ∑ i = 1 n ( x i − x ¯ ) 2

we can rewrite this using expectations (E[]):

LaTeX: var\left(x\right) = \frac{1}{n-1}\sum_{i=1}^{n}(x_i-E[x])^2 \\
var\left(x\right) = E\left[(x_i-E[x])^2\right] \\
var\left(x\right) = E\left[ x^2 -xE[x] -E[x]x + E[x]^2 \right] \\
var\left(x\right) = E[ x^2] -E[x]E[x] -E[x]E[x] + E[x]^2 \\
 var\left(x\right) = E[ x^2] -E[x]^2  \\ 
 
v a r ( x ) = 1 n − 1 ∑ i = 1 n ( x i − E [ x ] ) 2 v a r ( x ) = E [ ( x i − E [ x ] ) 2 ] v a r ( x ) = E [ x 2 − x E [ x ] − E [ x ] x + E [ x ] 2 ] v a r ( x ) = E [ x 2 ] − E [ x ] E [ x ] − E [ x ] E [ x ] + E [ x ] 2 v a r ( x ) = E [ x 2 ] − E [ x ] 2
```{r}
var(x)
sum(x^2 - mean(x)^2)/(length(x)-1)
```
Covariance  

LaTeX: cov\left(x,y\right) = \frac{1}{n-1}\sum_{i=1}^{n}(x_i-\bar{x})(y_i-\bar{y}) 

c o v ( x , y ) = 1 n − 1 ∑ i = 1 n ( x i − x ¯ ) ( y i − y ¯ )

we can rewrite this using expectations (E[]):

LaTeX: var\left(x\right) = \frac{1}{n-1}\sum_{i=1}^{n}(x_i-E[x])(y_i-E[y]) \\
var\left(x\right) = E\left[(x_i-E[x])(y_i-E[y])\right] \\
var\left(x\right) = E\left[ xy -xE[y] -E[x]y + E[x]E[y] \right] \\
var\left(x\right) = E[ xy] -E[x]E[y] -E[x]E[y] + E[x]E[y] \\
 var\left(x\right) = E[ xy] -E[x]E[y]  \\ 
 
v a r ( x ) = 1 n − 1 ∑ i = 1 n ( x i − E [ x ] ) ( y i − E [ y ] ) v a r ( x ) = E [ ( x i − E [ x ] ) ( y i − E [ y ] ) ] v a r ( x ) = E [ x y − x E [ y ] − E [ x ] y + E [ x ] E [ y ] ] v a r ( x ) = E [ x y ] − E [ x ] E [ y ] − E [ x ] E [ y ] + E [ x ] E [ y ] v a r ( x ) = E [ x y ] − E [ x ] E [ y ]

```{r}
cov(x,y)
sum(x*y - mean(x)*mean(y))/(length(x)-1)
# covariance and correlation as matrices
X <- matrix(cbind(x,y),ncol=2)

# covariance
cov(x,y)
cov(X)
var(X)
N <- dim(X)[1]
Xmean <- matrix(rep(colMeans(X),N),nrow=N, byrow=T)
Xc <- X - Xmean # a "centered" version of X
S <- t(Xc) %*% (Xc) /(N-1) # covariance via multiplying centered matrix w/ itself
S

# correlation 
cor(x,y)
cor(X)
Xsd <- matrix(rep(sqrt(diag(S)),N),nrow=N, byrow=T) # the sd of each column of X repeated N times
Xs <- Xc/Xsd # a "scaled" and centered version of X
(t(Xs) %*% Xs) / (N-1)
```

version control and projects
* saving your workspace as various types of projects (project, package, shiny webapp, various R+cpp formats, and RMarkdown website) via File->New Project 
* loading experimental code libraries with
```devtools::install_github("r-lib/devtools")```
# instead of
```install.packages("devtools")```
 
pairs plot
http://www.sthda.com/english/wiki/scatter-plot-matrices-r-base-graphs 
https://www.r-bloggers.com/scatterplot-matrices-pair-plots-with-cdata-and-ggplot2/ 
